---
title: "Make Choose All Table"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Make Choose All Table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidyREDCap)
```

# First Load REDCap Data into R
The `redcapAPI` package can be used to load data directly into R.  To learn more about it, take a look [here](https://github.com/nutterb/redcapAPI/wiki).  Normally the code to automatically pull data with an API includes a person's secret code "key".  Because I want to keep this hidden, I have hidden this API key in my user profile and the code below includes a call to `Sys.getenv()` to grab the key.  To learn more about working with APIs look [here](https://daattali.gitbooks.io/stat545-ubc-github-io/bit003_api-key-env-var.html). Also notice that the data is saved using the `saveRDS()` function. REDCap data loaded with the API has the variable labels added as an extra attribute.  To allow this vignette to run without sharing my secret key I have saved the data to the package website.

```{r getData, eval=FALSE}

rcon <- redcapAPI::redcapConnection(
  url = 'https://redcap.miami.edu/api/',
  token = Sys.getenv("NCI_API_Key")
)

redcap <- redcapAPI::exportRecords(rcon)

saveRDS(redcap, file = "redcap.rds")
```

```{r}
redcap <- readRDS(file = "./redcap.rds")
```

# The Problem 

REDCap exports a "choose all that apply" question into a series of similarly-named, binary indicator variables (i.e., the variables are equal to either "checked" or "unchecked"). It is desirable to have a consise table showing how often each option was chosen.

# The Solution

If you pass the `make_choose_all_table()` function the name of a REDCap export and the name of the _choose all that apply question_ question in REDCap, it will produce a consise frequency count table.


```{r showCode, eval=FALSE}
make_choose_all_table(redcap, "ingredients") %>% 
  knitr::kable()
```

### What ingredients do you currently crave?

```{r showOutput, echo=FALSE}
make_choose_all_table(redcap, "ingredients") %>% 
  knitr::kable()
```

