---
title: "Make a 'Choose One' Table"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Make a 'Choose One' Table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#"
)
```

```{r setup}
library(tidyREDCap)
```

</br>


# The Problem
It is often desirable to print variable labels above a summary table that shows the count of factor labels.  The labels that are exported on _choose all that apply_ questions include both the question and whichever response was chosen.  This redundancy is often unwanted, and the results are not presented professionally.

For example, in the Nacho Craving Index data, the first ingredient is "Chips". We see how R presents this information by simply printing the components of the `ingredients___1` column.
```{r}
redcap <- readRDS(file = "./redcap.rds")
redcap$ingredients___1
```

As we can see, this information is quite ugly, so we want to tabulate the results instead. However, if we use the simple `table()` function to clean up this information, we lose the original question and the answer label for `ingredients___1`.
```{r}
table(redcap$ingredients___1)
```
We no longer know what the question was, or which "select all" option this information represents.

### Aside: Loading REDCap Data into R
The `redcapAPI` package can be used to load data directly into R.  To learn more about it, take a look [here](https://github.com/nutterb/redcapAPI/wiki).  Normally the code to automatically pull data with an API includes a person's secret code "key".  Because I want to keep this hidden, I have hidden this API key in my user profile and the code below includes a call to `Sys.getenv()` to grab the key.  To learn more about working with APIs, look [here](https://daattali.gitbooks.io/stat545-ubc-github-io/bit003_api-key-env-var.html). Also notice that the data is saved using the `saveRDS()` function. REDCap data loaded with the API has the variable labels added as an extra attribute.  To allow this vignette to run without sharing my secret key, I have saved the data to the package website.

```{r getData, eval=FALSE}
rcon <- redcapAPI::redcapConnection(
  url = 'https://redcap.miami.edu/api/', 
  token = Sys.getenv("NCI_API_Key")
)

redcap <- redcapAPI::exportRecords(rcon)
```

</br>
</br>


# The Solution
The `make_choose_one_table()` function can be used with a factor variable to tabulate the response *while preserving the question and checked option context*.
```{r example_raw}
make_choose_one_table(redcap$ingredients___1) 
```

Further, this output can be molded into a publication-ready table with a single additional function call.
```{r example_pretty, results="asis"}
make_choose_one_table(redcap$ingredients___1) %>% 
  knitr::kable()
```

The `subset` option, if set to `TRUE`, will cause the function to remove the label's text and only show the response option (i.e., not repeat the "What ingredients do you currently crave?" question).
```{r example2, results="asis", }
make_choose_one_table(
  redcap$ingredients___2,
  subset = TRUE
) %>% 
  knitr::kable()
```

This function can also be used in an analysis pipeline with a data frame name and the name of the factor inside of that data frame. For example:
```{r example3,  results="asis", fig.align="left"}
redcap %>% 
  make_choose_one_table(ingredients___3) %>% 
  knitr::kable() 
```

