---
title: "Make a 'Choose All' Table"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Make a 'Choose All' Table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(tidyREDCap)
library(dplyr)
```

</br>


# The Problem 

REDCap exports a "choose all that apply" question into a series of similarly-named, binary indicator variables (i.e., the variables are equal to either "checked" or "unchecked"). For example, the following data represents a sample of responses to the Nacho Craving Index.
```{r}
redcap <- readRDS(file = "./redcap.rds")
redcap %>% 
  select(starts_with("ingredients___")) %>% 
  head()
```
It is desirable to have a concise table showing how often each option was chosen.

## Aside: Loading REDCap Data into R
See the [Import All Instruments from a REDCap Project](../doc/importInstruments.html) and [Importing from REDCap](../doc/useAPI.html) vignettes for details/information.

# The Solution

If you pass the `make_choose_all_table()` function, the name of a REDCap export, and the name of the _choose all that apply question_ question in REDCap, it will produce a concise frequency count table. 
```{r show_results}
make_choose_all_table(redcap, "ingredients") 
```

Similar to the `make_choose_one_table()` function, we can use this function inside an analysis pipeline. We can add the `kable()` call to make the table publication quality.
```{r show_results_pretty, results='asis'}
redcap %>% 
  make_choose_all_table("ingredients") %>% 
  knitr::kable()
```
