---
title: "Make a 'Choose One' Table"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Make a 'Choose One' Table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#"
)
```

```{r setup}
library(tidyREDCap)
```

</br>


# The Problem
It is often desirable to print variable labels above a summary table that shows the count of factor labels. The labels exported on _choose all that apply_ questions, including the question and whichever response was chosen. This redundancy is often unwanted, and the results are not presented professionally.

For example, in the race/ethnicity data, the first race option is "American Indian/Alaska Native". We see how R presents this information by simply printing the components of the `race___1` column.
```{r}
sample_data <- readRDS(file = "./sample_race_ethnicity.rds")
sample_data$race___1
```

As we can see, this information is quite ugly, so we want to tabulate the results instead. However, if we use the simple `table()` function to clean up this information, we lose the original question and the answer label for `race___1`.
```{r}
table(sample_data$race___1)
```
We no longer know what the question was or which "select all" option this information represents.

## Aside: Loading REDCap Data into R
See the [Import All Instruments from a REDCap Project](../doc/importInstruments.html) and [Importing from REDCap](../doc/useAPI.html) vignettes for details/information.

# The Solution
The `make_choose_one_table()` function can be used with a factor variable to tabulate the response *while preserving the question and checked option context*.
```{r example_raw}
make_choose_one_table(sample_data$race___1) 
```

Further, this output can be molded into a publication-ready table with a single additional function call.
```{r example_pretty, results="asis"}
make_choose_one_table(sample_data$race___1) |> 
  knitr::kable()
```

The `subset` option, if set to `TRUE,` will cause the function to remove the label's text and only show the response option (i.e., not repeat the question).
```{r example2, results="asis", }
make_choose_one_table(
  sample_data$race___2,
  subset = TRUE
) |> 
  knitr::kable()
```

This function can also be used in an analysis pipeline with a data frame name and the name of the factor inside that data frame. For example:
```{r example3,  results="asis", fig.align="left"}
sample_data |> 
  make_choose_one_table(race___3) |> 
  knitr::kable() 
```

